---
title: "CreditRisk_HM6"
author: "Hardik Maheshwari"
date: "November 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
install.packages("knitr")
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r results='hide', echo = FALSE, message=FALSE, warning=FALSE}
install.packages('latticeExtra')
install.packages('data.table')
install.packages('dplyr')
install.packages('ggplot2')
install.packages('gridExtra')
install.packages('purrr')
install.packages('tidyr')
install.packages('grid')
install.packages('gtable')
install.packages('haven')
install.packages('lubridate')
install.packages('zoo')
install.packages('survminer')
install.packages('survival')
install.packages('glmnet')
install.packages('broom')
install.packages('gains')
install.packages('ROCR')
install.packages('pROC')
#install.packages('DMwR')
install.packages('caret')
install.packages('e1071')
install.packages('class')

library(data.table)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(purrr)
library(tidyr)
library(latticeExtra)
library(grid)
library(gtable)
library(haven)
library(lubridate)
library(zoo)
library(survminer)
library(survival)
library(glmnet)
library(broom)
library(gains)
library(ROCR)
library(pROC)
#library(DMwR)
library(caret)
library(e1071)
library(class)
```


## setting directory and creating dataset

```{r results='hide', message=FALSE, echo=FALSE, warning=FALSE}
setwd("P:/MFI/assignment6")

dfLendingClub <-  read.csv("Q:/Data-ReadOnly/LendingClub/lc_loans_2007_2014.csv")

dfLendingClub <- dfLendingClub[,c('Default','issue_d','last_pymnt_d','loan_amnt','home_ownership','emp_length','acc_open_past_24mths','annual_inc','bc_open_to_buy','dti','inq_last_6mths','mort_acc','num_tl_op_past_12m','percent_bc_gt_75','tot_hi_cred_lim','total_bc_limit', 'tot_cur_bal', 'total_acc' , 'mo_sin_rcnt_tl' )]



#converting string variable to numeric variables
ownerLength <- unique(dfLendingClub$home_ownership)
dfLendingClub$home_ownership <- match(dfLendingClub$home_ownership, ownerLength)
empLength <- unique(dfLendingClub$emp_length)
dfLendingClub$emp_length <- match(dfLendingClub$emp_length, empLength)

#keep check of the na removal thing
dfLendingClub <- subset(dfLendingClub, !is.na(dfLendingClub$tot_hi_cred_lim))
dfLendingClub[is.na(dfLendingClub)] <- 0
dfLendingClub <- dfLendingClub[complete.cases(dfLendingClub),]

```

#calculation of descriptive statistics
```{r}
dfDescStats<- dfLendingClub[,c('Default','loan_amnt','home_ownership','emp_length','acc_open_past_24mths','annual_inc','bc_open_to_buy','dti','inq_last_6mths','mort_acc','num_tl_op_past_12m','percent_bc_gt_75','tot_hi_cred_lim','total_bc_limit','tot_cur_bal', 'total_acc' , 'mo_sin_rcnt_tl')]


pdf("desc_stats.pdf", width = 20, height = 30)
plot.new()
string <- "Descriptive Stats"
title(string, cex.main=2)
tmp <- do.call(data.frame, 
               list(count = apply(dfDescStats, 2, length),
                    mean = apply(dfDescStats, 2, mean),
                    Q_25 = apply(dfDescStats, 2, quantile, probs = c(0.25)),
                    Q_50 = apply(dfDescStats, 2, quantile, probs = c(0.50)),
                    Q_75 = apply(dfDescStats, 2, quantile, probs = c(0.75)),
                    sd = apply(dfDescStats, 2, sd),
                    min = apply(dfDescStats, 2, min),
                    max = apply(dfDescStats, 2, max)
               ))
grid.table(tmp)
dev.off() 
rm(tmp)
rm(dfDescStats)
```

#Survival Analysis Part 3 Kaplan-Meier Survival
```{r}
defSurvival <-  dfLendingClub

defSurvival$issue_d  <- paste ('01-', defSurvival$issue_d , collapse = NULL)
defSurvival$issue_d  <-  gsub(" ","",defSurvival$issue_d ) 
defSurvival$issueDate <-  as.Date(as.character(defSurvival$issue_d)  , format='%d-%b-%Y')
defSurvival$last_pymnt_d  <- paste ('01-', defSurvival$last_pymnt_d , collapse = NULL)
defSurvival$last_pymnt_d  <-  gsub(" ","",defSurvival$last_pymnt_d ) 
defSurvival$lastPaymentDate <-  as.Date(as.character(defSurvival$last_pymnt_d)  , format='%d-%b-%Y')
defSurvival$time <- interval(defSurvival$issueDate , defSurvival$lastPaymentDate) %/% months(1)

#charged off means status = 1
pdf("plotsPart3.pdf", width = 20, height = 30)
fit <- survfit(Surv(time, Default) ~ 1, data = defSurvival)
#summary(fit)
ggsurvplot(fit, data = defSurvival, risk.table = TRUE,  conf.int = TRUE, palette = "Dark2", 
           censor = FALSE, surv.median.line = "hv", title="Kaplan Meier Survival")
dev.off() 

#group wise based on homeownership as category
catvar <- c('MORTGAGE', 'RENT', 'OWN','ANY','OTHER','NONE' )
pdf("plotsPart3GroupWise.pdf", width = 20, height = 30)
ho <- unique(defSurvival$home_ownership)
fit_list <- list()
for (i in ho){
  smallF <- subset(defSurvival, defSurvival$home_ownership == i)
  if( nrow(smallF)> 10& i != 4){
    fit <- survfit(Surv(time, Default ) ~ Default, data = smallF)
    #summary(fit)
    fit_list[[i]] <- ggsurvplot(fit, data = smallF, risk.table = TRUE,  conf.int = TRUE, palette = "Dark2", 
           censor = FALSE, surv.median.line = "hv",title = paste ("Kaplan Meier Survival HOMEOWNERSHIP as Category Variable", catvar[i]))
  }
}  
arrange_ggsurvplots(fit_list[-4])
dev.off() 
rm(smallF)
```
# Survival Analysis using Cox Proportional Heads Model
```{r}

#charged off means status = 1
pdf("plotsPart4.pdf", width = 20, height = 30)
res.cox <- coxph(Surv(time, Default) ~ loan_amnt + home_ownership + emp_length + acc_open_past_24mths +annual_inc + bc_open_to_buy + dti + inq_last_6mths + mort_acc + num_tl_op_past_12m + percent_bc_gt_75+ tot_hi_cred_lim + total_bc_limit +  tot_cur_bal+ total_acc, data = defSurvival)
#summary(fit)
ggsurvplot(survfit(res.cox ), data = defSurvival, risk.table = TRUE, title="Cox Proportional Heads")
dev.off() 

#group wise based on homeownership as category
catvar <- c('MORTGAGE', 'RENT', 'OWN','ANY','OTHER','NONE' )
pdf("plotsPart4GroupWise.pdf", width = 20, height = 30)
ho <- unique(defSurvival$home_ownership)
fit_list <- list()
for (i in ho){
  smallF <- subset(defSurvival, defSurvival$home_ownership == i)
  if( nrow(smallF)> 10 & i != 4){
    res.cox <- coxph(Surv(time, Default) ~ loan_amnt + home_ownership + emp_length + acc_open_past_24mths +annual_inc + bc_open_to_buy + dti + inq_last_6mths + mort_acc + num_tl_op_past_12m + percent_bc_gt_75+ tot_hi_cred_lim + total_bc_limit+tot_cur_bal+ total_acc + mo_sin_rcnt_tl, data = smallF)
    #summary(fit)
    fit_list[[i]] <- ggsurvplot(survfit(res.cox ), data = smallF, risk.table = TRUE, title = paste ("Cox Proportional Heads Survival HOMEOWNERSHIP as Category Variable", catvar[i]))
  }
}  
arrange_ggsurvplots(fit_list[-4]) #combine
dev.off() 
rm(defSurvival)
rm(res.cox)
rm(fit_list)
```

#part 5 insample logistic Regression- Entire Sample of Loans issued
```{r}
dfPart5 <- dfLendingClub
pdf("LogisticRegressionPart5.pdf", width = 20, height = 30)
plot.new()
string <- "Logistic Regression Entire Sample Part5"
title(string, cex.main=2)
logitPart5 <- glm(Default ~ loan_amnt + home_ownership + emp_length + acc_open_past_24mths +annual_inc + bc_open_to_buy + dti + inq_last_6mths + mort_acc + num_tl_op_past_12m + percent_bc_gt_75+ tot_hi_cred_lim + total_bc_limit+ tot_cur_bal+ total_acc + mo_sin_rcnt_tl, data = dfPart5, family = "binomial")
temp <- tidy(logitPart5)
grid.table(temp)
dev.off()
rm(dfPart5)
rm(temp)
rm(logitPart5)
```

#part6 Out of the Sample Logistic Regression (2007-2013 training data, 2014 testing data)
```{r}
dfPart6 <- dfLendingClub
dfPart6$issue_d  <- paste ('01-', dfPart6$issue_d , collapse = NULL)
dfPart6$issue_d  <-  gsub(" ","",dfPart6$issue_d ) 
dfPart6$issueDate <-  as.Date(as.character(dfPart6$issue_d)  , format='%d-%b-%Y')
dfPart6$YEAR <- year(dfPart6$issueDate)
years <- c(2007, 2008, 2009, 2010, 2011, 2012, 2013)
trainData <- filter(dfPart6, YEAR %in% years)
testData <- filter(dfPart6, YEAR %in% 2014)
testData <- subset(testData, home_ownership != 4)

# fitting the model
pdf("LogisticRegressionPart6.pdf", width = 20, height = 30)
plot.new()
string <- "Logistic Regression Part 6"
title(string, cex.main=2)
logitPart6 <- glm(Default ~ loan_amnt + home_ownership + emp_length + acc_open_past_24mths +annual_inc + bc_open_to_buy + dti + inq_last_6mths + mort_acc + num_tl_op_past_12m + percent_bc_gt_75+ tot_hi_cred_lim + total_bc_limit+ tot_cur_bal+ total_acc + mo_sin_rcnt_tl, data = trainData, family = "binomial")
temp <- tidy(logitPart6)
grid.table(temp)
dev.off()

testData$probs<- predict(logitPart6,newdata=testData , type="response")
testData$decile <- as.integer(cut(testData$probs, breaks=quantile(testData$probs, seq(0,1,.1), include.lowest=T)))

pdf("Part6DefaultDeciles.pdf", width = 20, height = 30)
plot.new()
string <- "Logistic Regression Deciles"
title(string, cex.main=2)
totalDefaults <- testData %>% group_by(decile) %>% summarise(Frequency = sum(Default))
totalDefaults <- totalDefaults[complete.cases(totalDefaults),]
grid.table(totalDefaults)
roc <- roc(testData$Default,testData$probs)
plot.roc(roc, print.auc = T, grid = T, xlim = c(1,0)) 
ROCRpred<-prediction(testData$probs,testData$Default)
plot(performance(ROCRpred, measure = 'tpr', x.measure = 'fpr'))
perf <- performance(ROCRpred, measure = "auc")
auc <- perf@y.values[[1]]


KSStats <- function(pred,depvar){
  require("ROCR")
  p   <- prediction(as.numeric(pred),depvar)
  perf <- performance(p, "tpr", "fpr")
  ks <- max(attr(perf, "y.values")[[1]] - (attr(perf, "x.values")[[1]]))
  return(ks)
}
dev.off()
print(perf)
print(KSStats(testData$probs,testData$Default))
rm(logitPart6)
```

#part 8 Lasso Regression (2007-2013 training data, 2014 testing data)
```{r}

dfLendingClub <-  read.csv("Q:/Data-ReadOnly/LendingClub/lc_loans_2007_2014.csv")
ownerLength <- unique(dfLendingClub$home_ownership)
dfLendingClub$home_ownership <- match(dfLendingClub$home_ownership, ownerLength)
empLength <- unique(dfLendingClub$emp_length)
dfLendingClub$emp_length <- match(dfLendingClub$emp_length, empLength)


#keep check of the na removal thing
#dfLendingClub <- subset(dfLendingClub, !is.na(dfLendingClub$tot_hi_cred_lim))
dfLendingClub[is.na(dfLendingClub)] <- 0
dfLendingClub <- dfLendingClub[complete.cases(dfLendingClub),]


x <- subset( dfLendingClub, select = -dfLendingClub$Default )
x <- subset(x, select = -c(issue_d, last_pymnt_d, earliest_cr_line, title))
x <- as.matrix(data.frame(x))


cvfit = cv.glmnet(x, y=as.factor(dfLendingClub$Default), family = "binomial", type.measure = "class")
plot(cvfit)
optimalLambda <- cvfit$lambda.min
coef(cvfit, s = "lambda.min")
print(coef(cvfit, s = "lambda.min"))
#dfLassoBased <- dfLendingClub[,c('Default','acc_open_past_24mths','annual_inc','bc_open_to_buy','dti','inq_last_6mths','mort_acc','num_tl_op_past_12m','percent_bc_gt_75','tot_hi_cred_lim','total_bc_limit')]


dfLendingClub$issue_d  <- paste ('01-', dfLendingClub$issue_d , collapse = NULL)
dfLendingClub$issue_d  <-  gsub(" ","",dfLendingClub$issue_d ) 
dfLendingClub$issueDate <-  as.Date(as.character(dfLendingClub$issue_d)  , format='%d-%b-%Y')
dfLendingClub$YEAR <- year(dfLendingClub$issueDate)
years <- c(2007, 2008, 2009, 2010, 2011, 2012, 2013)
trainData <- filter(dfLendingClub, YEAR %in% years)
testData <- filter(dfLendingClub, YEAR %in% 2014)
testData <- subset(testData, home_ownership != 4)


pdf("LogLasso8.pdf", width = 20, height = 30)
plot.new()
string <- "Lasso Regression Part8 minimum lambda details"
title(string, cex.main=2)
print(coef(cvfit, s = "lambda.min"))
dev.off()

pdf("LogLasso8_post.pdf", width = 20, height = 30)
plot.new()
string <- "Post lasso"
title(string, cex.main=2)

logitPart8 <- glm(Default ~ loan_amnt + home_ownership+ acc_open_past_24mths +annual_inc + bc_open_to_buy + dti + inq_last_6mths + mort_acc + num_tl_op_past_12m + percent_bc_gt_75+ tot_hi_cred_lim + total_bc_limit+ tot_cur_bal+ total_acc, data = trainData,family = "binomial")
temp <- tidy(logitPart8)
grid.table(temp)
dev.off()

testData$probs<- predict(logitPart8,newdata=testData , type="response")
testData$decile <- as.integer(cut(testData$probs, breaks=quantile(testData$probs, seq(0,1,.1), include.lowest=T)))

pdf("Part8DefaultDeciles.pdf", width = 20, height = 30)
plot.new()
string <- "Lasso Regression Part8 Deciles and Misclassfication Error"
title(string, cex.main=2)
totalDefaults <- testData %>% group_by(decile) %>% summarise(Frequency = sum(Default))
totalDefaults <- totalDefaults[complete.cases(totalDefaults),]
grid.table(totalDefaults)
roc <- roc(testData$Default,testData$probs)
plot.roc(roc, print.auc = T, grid = T, xlim = c(1,0)) 
ROCRpred<-prediction(testData$probs,testData$Default)
plot(performance(ROCRpred, measure = 'tpr', x.measure = 'fpr'))
perf <- performance(ROCRpred, measure = "auc")
auc <- perf@y.values[[1]]


KSStats <- function(pred,depvar){
  require("ROCR")
  p   <- prediction(as.numeric(pred),depvar)
  perf <- performance(p, "tpr", "fpr")
  ks <- max(attr(perf, "y.values")[[1]] - (attr(perf, "x.values")[[1]]))
  return(ks)
}
dev.off()




testData$prediction <- ifelse(testData$probs > 0.25, 1,0)
testData <- testData[complete.cases(testData),]
#testData$prediction <- as.factor(testData$prediction)
confMatrix <- confusionMatrix(table(testData$prediction, testData$Default))
res<- confMatrix$table
misclassfication <- 1-sum(diag(res))/sum(res)
print("Post Lasso Logistic Regression Measures")
print("Performance")
print(perf)
print("KS Statistics")
print(KSStats(testData$probs,testData$Default))
print("Misclassification")
print(misclassfication)



```
#part 9 K Nearest Neighbour
```{r}

# Part 9
xvar <- trainData[,c('loan_amnt', 'acc_open_past_24mths', 'annual_inc', 'bc_open_to_buy' , 'dti' , 'inq_last_6mths' , 'mort_acc', 'num_tl_op_past_12m' , 'percent_bc_gt_75' , 'tot_hi_cred_lim', 'total_bc_limit', 'tot_cur_bal','total_acc')]

# Normalizing Numeric Variable
normalize <- function(x) {
  return ((x - mean(x)) / sd(x)) }
xvar <- as.data.frame(lapply(xvar, normalize))
xvar <- cbind(trainData$home_ownership, xvar)
setnames(xvar,"trainData$home_ownership","home_ownership" )
xvar$home_ownership <- as.numeric(xvar$home_ownership)

xVarTest <- testData[,c('loan_amnt' , 'acc_open_past_24mths', 'annual_inc', 'bc_open_to_buy' , 'dti' , 'inq_last_6mths' , 'mort_acc', 'num_tl_op_past_12m' , 'percent_bc_gt_75' , 'tot_hi_cred_lim', 'total_bc_limit', 'tot_cur_bal','total_acc')]
xVarTest <- as.data.frame(lapply(xVarTest, normalize))
xVarTest <- cbind(testData$home_ownership, xVarTest)
setnames(xVarTest,"testData$home_ownership","home_ownership")
xVarTest$home_ownership <- as.numeric(xVarTest$home_ownership)

labels <- trainData$Default
#knn_model <- knn(xvar, xVarTest, cl = labels, k = 2)
#knn_model <- data.frame(knn_model)
#testData$Default <- as.factor(testData$Default)


cfm_list <- list()
df1 <- data.frame()

pdf("Part9KNN.pdf", width = 20, height = 30)
plot.new()
string <- "KNN Part9 Different K Values"
title(string, cex.main=2)
for (i in 2:10){
  print(i)
  knn_model <- knn(xvar, xVarTest, cl = labels, k = i)
  knn_model <- data.frame(knn_model)
  confMat <- confusionMatrix(table(knn_model$knn_model, testData$Default))
  res <- confMat$table
  misclassfication <- 1-sum(diag(res))/sum(res)
  df <- data.frame(i,misclassfication)
  df1 <- rbind(df1,df)
}
grid.table(df1)
dev.off()

knn_model <- knn(xvar, xVarTest, cl = labels, k = 9)
knn_model <- data.frame(knn_model)
confMat <- confusionMatrix(table(knn_model$knn_model, loan_acc_test$Default))
res <- confMat$table
misclassfication <- 1-sum(diag(res))/sum(res)


print("The Value of K is 9 for which there is least misclassfication error")
print("The misclassification error:")
print(misclassfication)



```