
```{r setup, include=FALSE}
install.packages("knitr")
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
```

```{r results='hide', echo = FALSE, message=FALSE, warning=FALSE}
install.packages('latticeExtra')
install.packages('data.table')
install.packages('dplyr')
install.packages('ggplot2')
install.packages('gridExtra')
install.packages('purrr')
install.packages('tidyr')
install.packages('sas7bdat')
install.packages('grid')
install.packages('gtable')
install.packages('haven')
install.packages('lubridate')
install.packages('rootSolve')
install.packages('zoo')

library(data.table, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(gridExtra, quietly = TRUE)
library(purrr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(latticeExtra, quietly = TRUE)
library(grid, quietly = TRUE)
library(gtable, quietly = TRUE)
library(sas7bdat, quietly = TRUE)
library(haven, quietly = TRUE)
library(lubridate, quietly = TRUE)
library(rootSolve, quietly = TRUE)
library(zoo)

```


## creating dataset  CRSP

```{r results='hide', message=FALSE, echo=FALSE, warning=FALSE}
setwd("P:/MFI/assignment5/5.3")
dfCRSP <- setDT(read_sas("Q:/Data-ReadOnly/CRSP/dsf.sas7bdat", cols_only = c('CUSIP','DATE','PRC','SHROUT','RET')))
#dfCRSP <- dfCRSP[1:100000] #comment this bit
dfCRSP$SHROUT <- dfCRSP$SHROUT * 1000
dfCRSP$YEAR <- year(dfCRSP$DATE)
dfCRSP <- dfCRSP[complete.cases(dfCRSP),]
dfCRSP$MARKETCAP <- abs(dfCRSP$PRC) * dfCRSP$SHROUT
dfANNRET  <- dfCRSP%>% group_by(CUSIP, YEAR) %>% summarise(annret = exp(sum(log(1+RET))))
dfSTDDev  <- dfCRSP%>% group_by(CUSIP, YEAR) %>% summarise(sigmae = sd(RET) * sqrt(250))
dfCRSP <- merge(dfCRSP, dfANNRET, by = c('YEAR', 'CUSIP'))
dfCRSP <- merge(dfCRSP, dfSTDDev, by = c('YEAR', 'CUSIP'))
dfCRSP <- filter(dfCRSP, MARKETCAP > 0)
dfCRSP <- dfCRSP[,c('CUSIP', 'DATE','YEAR','annret', 'sigmae', 'MARKETCAP')]
dfCRSP <- dfCRSP %>% rename(ANNRET = annret,SIGMAE = sigmae)
rm(dfANNRET,dfSTDDev)

setorder(dfCRSP, "CUSIP","DATE")


dfFunda <- setDT(read_sas("Q:/Data-ReadOnly/COMP/funda.sas7bdat", cols_only = c('CUSIP','DATADATE','INDFMT','DATAFMT','POPSRC','FIC','CONSOL','FYEAR', 'DLC', 'DLTT')))
dfFunda <- subset(dfFunda, (INDFMT=='INDL'& DATAFMT=='STD' & POPSRC=='D' & FIC=='USA' & CONSOL=='C' & FYEAR >=1970 & FYEAR <= 2017))
#dfFunda <- dfFunda[1:100000] #comment this bit
dfFunda$CUSIP <- substr(dfFunda$CUSIP , 1, 8)
dfFunda$YEAR <- dfFunda$FYEAR
dfFunda[is.na(dfFunda)] <- 0
dfFunda$DLC <- dfFunda$DLC * 1000000
dfFunda$DLTT <-  dfFunda$DLTT * 1000000
dfFunda$FVD <- dfFunda$DLC + 0.5 * dfFunda$DLTT
dfFunda <- dfFunda[,c('CUSIP','YEAR','FVD')]
dfFunda <- filter(dfFunda, FVD > 0)
#dfFunda  <- dfFunda %>% group_by(CUSIP, YEAR) %>% summarise(FVD= first(FVD))
dfFunda<- dfFunda %>% group_by(.dots = c("CUSIP", "YEAR"))%>% summarize(FVD =first(FVD))
setorder(dfFunda, "CUSIP","YEAR")
#dfCRSP <- dfCRSP[order(dfCRSP$DATE, dfCRSP$CUSIP),]
#dfFunda <- dfFunda[order(dfFunda$YEAR, dfFunda$CUSIP),]
#dfCRSP <- merge(dfCRSP, dfFunda, by = c('YEAR', 'CUSIP'))
#dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% fill(FVD)
#dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% mutate(FVD=zoo::na.locf(FVD))
```
#merge step to create the intersecting dataframe of CRSP and FUNDA  
```{r}
#dfCRSP<- merge(dfCRSP, dfFunda, by = c('CUSIP','YEAR'), all.x = T)
#rm(dfFunda)

dfCRSP <- inner_join(dfCRSP, dfFunda, by = c('CUSIP','YEAR'))
#write_sas(dfCRSP,"Q:/Users/hmaheshwari6/chelsea.sas7bdat")
rm(dfFunda)

```

#Merging the dataset with interest rate data.
```{r}
#setwd("P:/MFI/assignment5/5.3")
#dfCRSP <- setDT(read_sas("Q:/Users/hmaheshwari6/chelsea.sas7bdat"))

rate <- read.csv("DAILYFED.csv")
rate$DATE <- as.Date(rate$DATE, "%Y-%m-%d")
rate <- data.table(rate)
rate$DTB3Cont <- log(1 +as.numeric(as.character(rate$DTB3)) * 0.01)
rate <- na.locf(na.locf(rate), fromLast= TRUE)
rate <- rate[,c('DATE', 'DTB3Cont')]
dfCRSP <-  inner_join(dfCRSP, rate , by = c('DATE')) #merge(dfCRSP, rate, by = c('DATE'))
rm(rate)
dfCRSP <- dfCRSP[,c('DATE', 'YEAR','CUSIP','SIGMAE','ANNRET','MARKETCAP','FVD', 'DTB3Cont')]
dfCRSP <- dfCRSP[complete.cases(dfCRSP),]
```

# Creating time series of the value of the assets for every CUSIP every Year
# using the value of assets and taking standard deviation to converge
```{r}
rootKMV <- function(x, parms) c(F1 = x*pnorm((log(x/parms[1])+ parms[2] + parms[3]^2/2)/parms[3])- exp(-parms[2])*parms[1]*pnorm((log(x/parms[1])+ parms[2] - parms[3]^2/2)/parms[3]) - parms[4])

findingSol = function(data){
  rootFrame <- data.frame()
  rootFrame <- data
  rootFrame$ValueF<- rep(-1,length=nrow(rootFrame))
  rootFrame$sigmav <- rootFrame$SIGMAE
  sigmaVF <- unique(rootFrame$sigmav)
  for (i in (1:nrow(rootFrame)) ){
    params = c(as.numeric(rootFrame$FVD[i]), as.numeric(rootFrame$DTB3Cont[i]), as.numeric(sigmaVF) ,   as.numeric(rootFrame$MARKETCAP[i]))
    
    rootFrame$ValueF[i] <- uniroot(rootKMV, parms = params,interval = c(-100000,100000000000), f.lower = -100000, f.upper = 100000000000 )$root
  }
  rootFrame[apply(rootFrame, 1, function(x) all(is.finite(x))), ]
  rootFrame <- rootFrame %>% mutate(vsolag = dplyr::lag(ValueF, n = 1, default = NA))
  rootFrame <- rootFrame[complete.cases(rootFrame),]
  rootFrame$valRet = (rootFrame$ValueF - rootFrame$vsolag)/ rootFrame$vsolag
  sigmaUP <- rootFrame%>% summarise(sigmae = sd(valRet) * sqrt(250))
  diff = abs(sigmaVF - sigmaUP)
  
  counter = 0
  while ((diff > 0.001) | (counter < 4)){
  #while(counter < 3){
    print(nrow(rootFrame))
    sigmaVF = sigmaUP
    if(nrow(rootFrame) <= 1){
       for (i in (1:nrow(rootFrame)) ){
           params = c(as.numeric(rootFrame$FVD[i]), as.numeric(rootFrame$DTB3Cont[i]), as.numeric(sigmaVF) ,   as.numeric(rootFrame$MARKETCAP[i]))
          rootFrame$ValueF[i] <- uniroot(rootKMV, parms = params,interval = c(-100000,100000000000), f.lower = -100000, f.upper = 100000000000)$root
        }
        rootFrame[apply(rootFrame, 1, function(x) all(is.finite(x))), ]
        rootFrame <- rootFrame %>% mutate(vsolag = dplyr::lag(ValueF, n = 1, default = NA))
        rootFrame <- rootFrame[complete.cases(rootFrame),]
        rootFrame$valRet = (rootFrame$ValueF - rootFrame$vsolag)/ rootFrame$vsolag
        sigmaUP <- rootFrame%>% summarise(sigmae = sd(valRet) * sqrt(250))
        diff = abs(sigmaVF - sigmaUP)
        counter = counter + 1
      }
      #print("chelsea")
    else{
      return (rootFrame)
    }
      #rootFrame[apply(rootFrame, 1, function(x) all(is.finite(x))), ]
      return (rootFrame)
  }
   
}

dfCRSPF <- dfCRSP%>% group_by(CUSIP, YEAR) %>%  do(data.frame(findingSol(.)))
#write_sas(dfCRSPF,"Q:/Users/hmaheshwari6/chelseaFinal.sas7bdat")
#dfCRSP <- dfCRSP[complete.cases(dfCRSP),]
rm(dfCRSP)
dfCRSP <- dfCRSPF
rm(dfCRSPF)
```

#lagging the variables 
```{r}
dfCRSP <- setDT(read_sas("Q:/Users/hmaheshwari6/chelseaFinal.sas7bdat"))
dfCRSP <-  dfCRSP[,c('DATE', 'YEAR','CUSIP','SIGMAE','ANNRET','MARKETCAP','FVD', 'DTB3Cont', 'ValueF', 'sigmav')]
dfCRSP <- dfCRSP %>% rename(ValueFirm = ValueF)
dfCRSP <- dfCRSP %>% rename(SIGMAV = sigmav)

dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% mutate(ANNRET = dplyr::lag(ANNRET, n = 1, default = NA))
dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% mutate(SIGMAE = dplyr::lag(SIGMAE, n = 1, default = NA))
dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% mutate(FVD = dplyr::lag(FVD, n = 1, default = NA))
dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% mutate(MARKETCAP = dplyr::lag(MARKETCAP, n = 1, default = NA))
dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% mutate(ValueFirm = dplyr::lag(ValueFirm, n = 1, default = NA))
dfCRSP <- dfCRSP %>% group_by(CUSIP) %>% mutate(SIGMAV = dplyr::lag(SIGMAV, n = 1, default = NA))
dfCRSP <- dfCRSP[complete.cases(dfCRSP),]
dfCRSP <- dfCRSP[order(dfCRSP$DATE, dfCRSP$CUSIP),]

dfCRSP  <- dfCRSP %>% group_by(YEAR, CUSIP) %>% summarise(rfrate= first(DTB3Cont),
                                                   FVD = first(FVD),
                                                   MARKETCAP =first(MARKETCAP),
                                                   ANNRET = first(ANNRET),
                                                   SIGMAE = first(SIGMAE),
                                                   ValueFirm = first(ValueFirm),
                                                   SIGMAV = first(SIGMAV)
                                                   )
```

#calculation of distance to default and probability of default for each firms for each year
```{r}

pdf("DDPDkmvValues.pdf", width = 20, height = 30)
plot.new()
dfCRSP$DDkmv <- (log(dfCRSP$ValueFirm/dfCRSP$FVD) + (dfCRSP$ANNRET - 0.5 * dfCRSP$SIGMAV* dfCRSP$SIGMAV) )/(dfCRSP$SIGMAV)
dfCRSP$PDkmv<- pnorm(-dfCRSP$DDkmv)
temp <- dfCRSP[,c('YEAR','CUSIP','DDkmv','PDkmv')]
years <- unique(temp$YEAR)
for (i in years){
   grid.newpage()
   temp2 <- subset(temp, temp$YEAR == i)
   grid.table(temp2)
}
dev.off()  
rm(temp)


```
#compute descriptive stats
```{r}

dfCRSP <- filter(dfCRSP, DDkmv < 1000)
dfCRSP <- filter(dfCRSP, DDkmv > 0)
ddpdFinal <- dfCRSP[,c('DDkmv','PDkmv')]

pdf("desc_stats.pdf", width = 20, height = 30)
  plot.new()
  string <- "Descriptive Stats"
  title(string, cex.main=2)
  tmp <- do.call(data.frame, 
                 list(count = apply(ddpdFinal, 2, length),
                      mean = apply(ddpdFinal, 2, mean),
                      Q_25 = apply(ddpdFinal, 2, quantile, probs = c(0.25)),
                      Q_50 = apply(ddpdFinal, 2, quantile, probs = c(0.50)),
                      Q_75 = apply(ddpdFinal, 2, quantile, probs = c(0.75)),
                      sd = apply(ddpdFinal, 2, sd)
                 ))
  grid.table(tmp)
dev.off()  
rm(tmp)
rm(ddpdFinal)
```
# correlation between method ddnaive, method direct & method iterative

```{r}
#naive method
dfCRSP$sigNaive1= (dfCRSP$MARKETCAP/(dfCRSP$MARKETCAP+dfCRSP$FVD)) * dfCRSP$SIGMAE  + (dfCRSP$FVD/(dfCRSP$MARKETCAP+dfCRSP$FVD)) * (dfCRSP$SIGMAE * 0.25 + 0.05)
dfCRSP$DDNaive = (log((dfCRSP$MARKETCAP+dfCRSP$FVD)/dfCRSP$FVD) + (dfCRSP$ANNRET - 0.5 * dfCRSP$sigNaive1* dfCRSP$sigNaive1) )/(dfCRSP$sigNaive1)


dfCRSP <- dfCRSP[complete.cases(dfCRSP),]
dfCRSP$PDNaive<- pnorm(-dfCRSP$DDNaive)

#direct method
calModel <- function(x) {
F1 = E - x[1]*pnorm( (log(x[1]/F)+(r+x[2]^2/2))/x[2] ) + F*exp(-r)*pnorm( (log(x[1]/F)+(r+x[2]^2/2))/x[2]-x[2])
F2 = sigmaE - x[1]/E * pnorm( (log(x[1]/F)+(r+x[2]^2/2))/x[2])* x[2]
c(F1 = F1, F2 = F2)
}

rootFrame <- select(dfCRSP,YEAR,CUSIP)
rootFrame$ValueF<- rep(-1,length=nrow(rootFrame))
rootFrame$sigmaVal<- rep(-1,length=nrow(rootFrame))


for (i in (1:nrow(dfCRSP)) ){
    E = dfCRSP$MARKETCAP[i]
    F = dfCRSP$FVD[i]
    r = dfCRSP$rfrate[i]
    sigmaE = dfCRSP$SIGMAE[i]
    solution <- multiroot(f = calModel, start = c(dfCRSP$MARKETCAP[i]+dfCRSP$FVD[i],dfCRSP$MARKETCAP[i]*dfCRSP$sigNaive1[i]/(dfCRSP$MARKETCAP[i]+dfCRSP$FVD[i])))
    rootFrame$ValueF[i] = solution$root[1]
    rootFrame$sigmaVal[i] = solution$root[2]
}

dfCRSP$DDDirect <- ( log(rootFrame$ValueF/dfCRSP$FVD) + (dfCRSP$ANNRET-rootFrame$sigmaVal**2/2)) / (rootFrame$sigmaVal)
dfCRSP$PDDirect <- pnorm(-dfCRSP$DDDirect)

rm(rootFrame)
dfCRSP <- dfCRSP[complete.cases(dfCRSP),]
ddpdFinal <- dfCRSP[,c('DDkmv','DDDirect', 'DDNaive')]

pdf("CorrelationDD.pdf", width = 20, height = 30)
  plot.new()
  string <- "Correlation Distance to Default"
  title(string, cex.main=2)
  tmp <- cor(ddpdFinal)
  grid.table(tmp)
dev.off()

ddpdFinal <- dfCRSP[,c('PDkmv','PDDirect', 'PDNaive')]

pdf("CorrelationPD.pdf", width = 20, height = 30)
  plot.new()
  string <- "Correlation Probability of Default"
  title(string, cex.main=2)
  tmp <- cor(ddpdFinal)
  grid.table(tmp)
dev.off()  
rm(tmp)
rm(ddpdFinal)
```
#Plot Mean, 25th 50th 75th Percentile of DD for iterative method, Direct Method, Naive Method
```{r}

pdf("plotsmean.pdf", width = 20, height = 30)
fmean <- function(df, columnVar, headerLeft) {
  df[ c( columnVar ,  "YEAR" )  ]  %>% group_by(YEAR) %>% summarise_(valueCol = paste( "mean(" ,  as.name(columnVar), ")"  ) ) %>%  ggplot(aes(x = YEAR, y = valueCol)) + geom_line() + labs(y = headerLeft) + ggtitle("Mean Plot")
}
namVar <- c('mean_DD_KMV_IterativeMethod', 'mean_DD_DirectMethod', 'mean_DD_NaiveMethod')
            
items <- c("DDkmv","DDDirect", "DDNaive")
K <- 1
for(i in 1:length(items)){
    plot(fmean(dfCRSP, items[[i]], namVar[[K]]))
    K <- K + 1
}
dev.off()

```
#Plot Quantiles of DD for iterative method, Direct Method, Naive Method
```{r}

pdf("quantiles.pdf", width = 20, height = 30)
fQuant <- function(df, probQuant, columnVar, headerLeft) {
  df[ c( columnVar ,  "YEAR" )  ]  %>% group_by(YEAR) %>% summarise_(valueCol = paste( "quantile(" ,  as.name(columnVar), ",probs=",probQuant ,")"  ) ) %>%  ggplot(aes(x = YEAR, y = valueCol)) + geom_line() + labs(y = headerLeft) + ggtitle("Quantile Plot")
}
setorder(dfCRSP, 'YEAR')

namVar <- c('Q_25_DD_KMV_IterativeMethod', 'Q_50_DD_KMV_IterativeMethod', 'Q_75_DD_KMV_IterativeMethod',
            'Q_25_DD_DirectMethod', 'Q_50_DD_DirectMethod', 'Q_75_DD_DirectMethod',
            'Q_25_DD_NaiveMethod', 'Q_50_DD_NaiveMethod', 'Q_75_DD_NaiveMethod'
            )

items <- c("DDkmv","DDDirect", "DDNaive")
quant <- c(0.25, 0.5, 0.75)
K <- 1
for(i in 1:length(items)){
  for(j in 1:length(quant)){
    plot(fQuant(dfCRSP, quant[[j]], items[[i]], namVar[[K]]))
    K <- K + 1
  }
}

dev.off()
```