---
title: "Maheshwari_Hardik_5_2"
author: "Hardik"
date: "October 21, 2018"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
install.packages("knitr")
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

```{r results='hide', echo = FALSE, message=FALSE, warning=FALSE}
install.packages('latticeExtra')
install.packages('data.table')
install.packages('dplyr')
install.packages('ggplot2')
install.packages('gridExtra')
install.packages('purrr')
install.packages('tidyr')
install.packages('sas7bdat')
install.packages('grid')
install.packages('gtable')
install.packages('haven')
install.packages('lubridate')
install.packages('rootSolve')
library(data.table, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(gridExtra, quietly = TRUE)
library(purrr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(latticeExtra, quietly = TRUE)
library(grid, quietly = TRUE)
library(gtable, quietly = TRUE)
library(sas7bdat, quietly = TRUE)
library(haven, quietly = TRUE)
library(lubridate, quietly = TRUE)
library(rootSolve, quietly = TRUE)
```


## creating dataset  CRSP

```{r results='hide', message=FALSE, echo=FALSE, warning=FALSE}
setwd("P:/MFI/assignment5/5.2")
dfCRSP <- setDT(read_sas("Q:/Data-ReadOnly/CRSP/dsf.sas7bdat", cols_only = c('CUSIP','DATE','PRC','SHROUT','PRC','RET')))
#dfCRSP <- dfCRSP[1:100000] #comment this bit
dfCRSP$SHROUT <- dfCRSP$SHROUT * 1000
dfCRSP$YEAR <- year(dfCRSP$DATE)
dfCRSP <- dfCRSP[complete.cases(dfCRSP),]
dfCRSP$MARKETCAP <- abs(dfCRSP$PRC) * dfCRSP$SHROUT
dfCRSP_Year  <- dfCRSP%>% group_by(CUSIP, YEAR) %>% summarise(annret = exp(sum(log(1+RET))), sigmae = sd(RET) * sqrt(250), MARKETCAP = first(MARKETCAP))
dfCRSP_Year <- dfCRSP_Year %>% group_by(CUSIP) %>% mutate(annretLag = dplyr::lag(annret, n = 1, default = NA))
dfCRSP_Year <- dfCRSP_Year %>% group_by(CUSIP) %>% mutate(sdLag = dplyr::lag(sigmae, n = 1, default = NA))
dfCRSP_Year <- dfCRSP_Year[complete.cases(dfCRSP_Year),]
dfCRSP_Year <- dfCRSP_Year[,c('CUSIP','YEAR','annretLag', 'sdLag', 'MARKETCAP')]
dfCRSP_Year <- dfCRSP_Year %>% rename(ANNRET = annretLag,SIGMAE = sdLag)

```

## creating funda dataset 
```{r results='hide', echo=FALSE, message=FALSE, warning=FALSE}
dfFunda <- setDT(read_sas("Q:/Data-ReadOnly/COMP/funda.sas7bdat", cols_only = c('CUSIP','INDFMT','DATAFMT','POPSRC','FIC','CONSOL','FYEAR', 'DLC', 'DLTT')))
dfFunda <- subset(dfFunda, (INDFMT=='INDL'& DATAFMT=='STD' & POPSRC=='D' & FIC=='USA' & CONSOL=='C' & FYEAR >=1970 & FYEAR <= 2017))
#dfFunda <- dfFunda[1:100000] #comment this bit
dfFunda$CUSIP <- substr(dfFunda$CUSIP , 1, 8)
dfFunda$YEAR <- dfFunda$FYEAR
dfFunda[is.na(dfFunda)] <- 0
dfFunda$DLC <- dfFunda$DLC * 1000000
dfFunda$DLTT <-  dfFunda$DLTT * 1000000
dfFunda$FVD <- dfFunda$DLC + 0.5 * dfFunda$DLTT
dfFunda <- dfFunda[,c('CUSIP','YEAR','FVD')]
dfFunda <- dfFunda %>% group_by(CUSIP) %>% mutate(FVDLAG = dplyr::lag(FVD, n = 1, default = NA))
dfFunda <- dfFunda[complete.cases(dfFunda),]
dfFunda <- dfFunda[,c('CUSIP','YEAR','FVDLAG')]
dfFunda <- dfFunda %>% rename(FVD = FVDLAG)
```
##merging the two dataset
```{r results='hide', message=FALSE, warning=FALSE}
dfFunda <- filter(dfFunda, FVD > 0)
dfDDNaive <- merge(dfCRSP_Year, dfFunda, by = c('CUSIP','YEAR'))
dfDDNaive[order(dfDDNaive$YEAR), ]
#Approach 1 -> 0.05 + 0.25 * sigma 
dfDDNaive$sigNaive1= (dfDDNaive$MARKETCAP/(dfDDNaive$MARKETCAP+dfDDNaive$FVD)) * dfDDNaive$SIGMAE  + (dfDDNaive$FVD/(dfDDNaive$MARKETCAP+dfDDNaive$FVD)) * (dfDDNaive$SIGMAE * 0.25 + 0.05)
dfDDNaive$distDefault1 = (log((dfDDNaive$MARKETCAP+dfDDNaive$FVD)/dfDDNaive$FVD) + (dfDDNaive$ANNRET - 0.5 * dfDDNaive$sigNaive1* dfDDNaive$sigNaive1) )/(dfDDNaive$sigNaive1)


dfDDNaive <- dfDDNaive[complete.cases(dfDDNaive),]
dfDDNaive$pdef1 <- pnorm(-dfDDNaive$distDefault1)
```

## rate dataset
```{r}
rate <- read.csv("DAILYFED.csv")
rate$DATE <- as.Date(rate$DATE, "%Y-%m-%d")
rate <- data.table(rate)
rate$DTB3Cont <- log(1 +as.numeric(as.character(rate$DTB3)) * 0.01)
rate[,YEAR := year(DATE)]
rate <- rate[complete.cases(rate),]
rate  <- rate %>% group_by(YEAR) %>% summarise(r= first(DTB3Cont))
```
#merge finalframe with rate
```{r}
dfDDNaive <- merge(dfDDNaive, rate, by = c('YEAR'))
dfDDNaive[order(dfDDNaive$YEAR), ]

```
#root solve direct approach for calculating v and sigmav
```{r}

calModel <- function(x) {
F1 = E - x[1]*pnorm( (log(x[1]/F)+(r+x[2]^2/2))/x[2] ) + F*exp(-r)*pnorm( (log(x[1]/F)+(r+x[2]^2/2))/x[2]-x[2])
F2 = sigmaE - x[1]/E * pnorm( (log(x[1]/F)+(r+x[2]^2/2))/x[2])* x[2]
c(F1 = F1, F2 = F2)
}

rootFrame <- select(dfDDNaive,YEAR,CUSIP)
rootFrame$ValueF<- rep(-1,length=nrow(rootFrame))
rootFrame$sigmaVal<- rep(-1,length=nrow(rootFrame))


for (i in (1:nrow(dfDDNaive)) ){
    E = dfDDNaive$MARKETCAP[i]
    F = dfDDNaive$FVD[i]
    r = dfDDNaive$r[i]
    sigmaE = dfDDNaive$SIGMAE[i]
    solution <- multiroot(f = calModel, start = c(dfDDNaive$MARKETCAP[i]+dfDDNaive$FVD[i],dfDDNaive$MARKETCAP[i]*dfDDNaive$sigNaive1[i]/(dfDDNaive$MARKETCAP[i]+dfDDNaive$FVD[i])))
    rootFrame$ValueF[i] = solution$root[1]
    rootFrame$sigmaVal[i] = solution$root[2]
}

dfDDNaive$distDefDirect <- ( log(rootFrame$ValueF/dfDDNaive$FVD) + (dfDDNaive$ANNRET-rootFrame$sigmaVal**2/2)) / (rootFrame$sigmaVal)
dfDDNaive$pdefDirect <- pnorm(-dfDDNaive$distDefDirect)


```
# Printing descriptive stats
```{r fig, fig.height = 15, fig.width = 15, out.width = '40%',  fig.align = "center"}
dfDDNaive <- dfDDNaive[complete.cases(dfDDNaive),]
dfDDNaive <- filter(dfDDNaive, distDefDirect < 1000)
dfDDNaive <- filter(dfDDNaive, distDefDirect > 0)
ddpdFinal <- dfDDNaive[,c('distDefDirect','pdefDirect')]
#part1 <- dfDDNaive[,c('distDefault1','pdef1','distDefault2','pdef2', 'distDefault3', 'pdef3')]

pdf("desc_stats.pdf", width = 20, height = 30)
  plot.new()
  string <- "Descriptive Stats"
  title(string, cex.main=2)
  tmp <- do.call(data.frame, 
                 list(count = apply(ddpdFinal, 2, length),
                      mean = apply(ddpdFinal, 2, mean),
                      Q_25 = apply(ddpdFinal, 2, quantile, probs = c(0.25)),
                      Q_50 = apply(ddpdFinal, 2, quantile, probs = c(0.50)),
                      Q_75 = apply(ddpdFinal, 2, quantile, probs = c(0.75)),
                      sd = apply(ddpdFinal, 2, sd)
                 ))
  grid.table(tmp)
dev.off()  
#rm(part1)
#rm(tmp)
#rm(string)
```
#computing correlation
```{r , results='asis', message=FALSE, warning=FALSE}
pdf("correlation.pdf", width = 20, height = 30)
plot.new()
  string <- "Correlation using Direct Approach Method"
  title(string, cex.main=2)
corrApproach1 <- cor(dfDDNaive$distDefDirect, dfDDNaive$pdefDirect)
grid.table(corrApproach1)
dev.off() 
```

#Plot for means
```{r}
pdf("plotsmean.pdf", width = 20, height = 30)
fmean <- function(df, columnVar, headerLeft) {
  df[ c( columnVar ,  "YEAR" )  ]  %>% group_by(YEAR) %>% summarise_(valueCol = paste( "mean(" ,  as.name(columnVar), ")"  ) ) %>%  ggplot(aes(x = YEAR, y = valueCol)) + geom_line() + labs(y = headerLeft) + ggtitle("Mean Plot")
}
namVar <- c('mean_DD_DirectApproach', 'mean_PD_DirectApproach')
            
items <- c("distDefDirect","pdefDirect")
K <- 1
for(i in 1:length(items)){
    plot(fmean(dfDDNaive, items[[i]], namVar[[K]]))
    K <- K + 1
}
dev.off()
```





#plot for quantiles
```{r}
pdf("quantiles.pdf", width = 20, height = 30)
fQuant <- function(df, probQuant, columnVar, headerLeft) {
  df[ c( columnVar ,  "YEAR" )  ]  %>% group_by(YEAR) %>% summarise_(valueCol = paste( "quantile(" ,  as.name(columnVar), ",probs=",probQuant ,")"  ) ) %>%  ggplot(aes(x = YEAR, y = valueCol)) + geom_line() + labs(y = headerLeft) + ggtitle("Quantile Plot")
}
setorder(dfDDNaive, 'YEAR')
#df <- dfDDNaive[,:= mean(distDefault1), by = c('YEAR')]
namVar <- c('Q_25_DD_DirectApproach', 'Q_50_DirectApproach', 'Q_75_DirectApproach',
            'Q_25_PD_DirectApproach', 'Q_50_PD_DirectApproach', 'Q_75_PD_DirectApproach')

items <- c("distDefDirect","pdefDirect")
quant <- c(0.25, 0.5, 0.75)
K <- 1
for(i in 1:length(items)){
  for(j in 1:length(quant)){
    plot(fQuant(dfDDNaive, quant[[j]], items[[i]], namVar[[K]]))
    K <- K + 1
  }
}

dev.off()
```

#NBER Recession Data Frame Generation
```{r}

usrec <- read.csv("USREC.csv")
usrec$DATE <- as.Date(usrec$DATE, "%Y-%m-%d")
usrec <- data.table(usrec)
usrec[,YEAR := year(DATE)]
usrec[,month := month(DATE)]
usrec <- usrec[YEAR %in% unique(YEAR) & month == 12]
usrec <- usrec[,c('YEAR','USREC')]
usrec <- unique(usrec)
dfForRecession <- dfDDNaive[,c('YEAR', 'distDefDirect','pdefDirect')]
dfForRecFinal <- aggregate(dfForRecession[, 1:3], list(dfForRecession$YEAR), mean)
dfForRecFinal <- dfForRecFinal[,c('YEAR', 'distDefDirect','pdefDirect')]
dfRecMerged <- merge(dfForRecFinal, usrec, by = c('YEAR'))
```
#NBER Recession Data Descriptive Stats
```{r fig, fig.height = 15, fig.width = 20, out.width = '40%',  fig.align = "center"}
pdf("nber_descstats.pdf", width = 20, height = 30)
NB <- unique(dfRecMerged$USREC)
for (i in NB){
  plot.new()
  string <- paste ("Descriptive Stats for USREC Value", i)
  title(string, cex.main=2)
  smallF <- subset(dfRecMerged, dfRecMerged$USREC == i)
  smallF <- smallF[, 2:3] 
  tmp <- do.call(data.frame, list(mean = apply(smallF, 2, mean),
                                  sd = apply(smallF, 2, sd),
                                  median = apply(smallF, 2, median),
                                  min = apply(smallF, 2, min),
                                  max = apply(smallF, 2, max),
                                  count = apply(smallF, 2, length),
                                  Q_25 = apply(smallF, 2, quantile, probs = c(0.25)),
                                  Q_50 = apply(smallF, 2, quantile, probs = c(0.50)),
                                  Q_75 = apply(smallF, 2, quantile, probs = c(0.75))
                                  )
                 )
 grid.table(tmp)
}
dev.off()
```
#NBER Recession Plot
```{R}
pdf("nberRecessionPlots.pdf", width = 20, height = 30)
NB <- unique(dfRecMerged$USREC)
for (i in NB){
  smallF <- subset(dfRecMerged, dfRecMerged$USREC == i)
  smallF <- smallF[, 1:3]
  cols <- colnames(smallF)
  cols <- cols[2:3]
  for (j in cols){
    plot_graph <- smallF[,c('YEAR',j)]
    plot(ggplot(plot_graph, aes_string(y = plot_graph[,c(2)], x = plot_graph[,c(1)]))+geom_line()+labs(x = "Date", y = colnames(plot_graph)[2])+ ggtitle(paste("USREC Value", i, " Plot for", colnames(plot_graph)[2])))
  }
}
dev.off()
```
#Moody's BAA-Fed Fund Spread
```{r}
moodyFrame <- read.csv("BAAFFM.csv")
moodyFrame$DATE <- as.Date(moodyFrame$DATE, "%Y-%m-%d")
moodyFrame <- data.table(moodyFrame)
moodyFrame$YEAR <- year(moodyFrame$DATE)
moodyFrame[,baa_avg := mean(BAAFFM), by = c('YEAR')]
moodyFrame <- moodyFrame[,c('YEAR','baa_avg')]
moodyFrame<- unique(moodyFrame)

dfForBAAFM <- dfDDNaive[,c('YEAR', 'distDefDirect','pdefDirect')]
dfForBAAFM <- aggregate(dfForBAAFM[, 1:3], list(dfForBAAFM$YEAR), mean)
dfForBAAFM <- dfForBAAFM[,c('YEAR', 'distDefDirect','pdefDirect')]
dfRecMergBAAFM <- merge(dfForBAAFM, moodyFrame, by = c('YEAR'))

pdf("moody.pdf", width = 20, height = 30)
cols <- colnames(dfRecMergBAAFM)
cols <- cols[2:3]
for (j in cols){
    plot_graph <- dfRecMergBAAFM[,c('YEAR',j, 'baa_avg')]
    plotname <-  paste("plot for", colnames(plot_graph)[3], "(Blue) vs", colnames(plot_graph)[2], " (Red)")
    p <- ggplot() +
      # blue plot
      geom_smooth(data=plot_graph, aes(x=plot_graph[,c(1)], y=plot_graph[,c(3)]), fill="blue",
        colour="darkblue", size=1) +
      # red plot
      geom_smooth(data=plot_graph, aes(x=plot_graph[,c(1)], y=plot_graph[,c(2)]), fill="red",
        colour="red", size=1) +
      labs(title=plotname, y="Values", x="Year")
    plot(p)
  }                                                                                                     

dev.off()
```
#Cleveland Stress Index
```{r}
CFSIndex <- read.csv("CFSI.csv")
CFSIndex <- data.table(CFSIndex)
CFSIndex$DATE <- as.Date(CFSIndex$DATE, "%Y-%m-%d")
CFSIndex[,YEAR := year(DATE)]
CFSIndex[,CFSIndex_avg := mean(CFSI), by = c('YEAR')]
CFSIndex <- CFSIndex[,c('YEAR','CFSIndex_avg')]
CFSIndex <- unique(CFSIndex)

dfForCSI <- dfDDNaive[,c('YEAR', 'distDefDirect','pdefDirect')]
dfForCSI <- aggregate(dfForCSI[, 1:3], list(dfForCSI$YEAR), mean)
dfForCSI <- dfForCSI[,c('YEAR', 'distDefDirect','pdefDirect')]
dfRecMergCFSI <- merge(dfForCSI, CFSIndex, by = c('YEAR'))

cols <- colnames(dfRecMergCFSI)
cols <- cols[2:3]
pdf("csi.pdf", width = 20, height = 30)
for (j in cols){
    plot_graph <- dfRecMergCFSI[,c('YEAR',j, 'CFSIndex_avg')]
    plotname <-  paste("plot for", colnames(plot_graph)[3], "(Blue) vs", colnames(plot_graph)[2], " (Red)")
    p <- ggplot() +
      # blue plot
      geom_smooth(data=plot_graph, aes(x=plot_graph[,c(1)], y=plot_graph[,c(3)]), fill="blue",
        colour="darkblue", size=1) +
      # red plot
      geom_smooth(data=plot_graph, aes(x=plot_graph[,c(1)], y=plot_graph[,c(2)]), fill="red",
        colour="red", size=1) +
      labs(title=plotname, y="Values", x="Year")
    plot(p)
}

dev.off()
```


